\chapter{Temperature Dependence of Magnetic Permeability\label{chap:muofT}}
This chapter is based on a study of the sensitivity of fields
generated within magnetically shielded volumes to changes in magnetic
permeability presented in Ref.~\cite{Andalib:2016ahj}.

The previous best nEDM measurement~\cite{Baker2006, Pendlebury2015} showed that
the dominant systematic uncertainty is related to the magnetic field
homogeneity and instability. As a result, much of the research and
development efforts for these experiments are focused on careful
design and testing of various magnetic shield geometries with
precision magnetometers~\cite{altarev2014magnetically,
  brys2005magnetic,afach2014dynamic,patton2014all}. As an example,
mechanical and temperature changes of the passive magnetic
shielding~\cite{voigt2013,thiel2007demagnetization}, and the
degaussing
procedure~\cite{thiel2007demagnetization,altarev2015minimizing,sun2016dynamic}
(also known as demagnetization, equilibration, or idealization),
affect the stability of the magnetic field within magnetically
shielded rooms. Using a shield-coupled coil, since the $B_0$ coil is
magnetically coupled to the innermost magnetic shield, any change in
the magnetic properties of the innermost shield will result in a change in
the internal magnetic field. This chapter focuses on this
effect and characterize one possible source of instability: changes of
the magnetic permeability $\mu$ of the material with temperature when
the magnetic shield is used as a return yoke for the internal coil
system. This shield-coupled coil design was used in the previous best
nEDM experiment.



%In the previous best experiment to measure
%nEDM~\cite{Baker2006,Pendlebury2015} which discovered
%$d_n<3.0\times 10^{-26}$~e$\cdot$cm~(90\% C.L), effects related to
%magnetic field homogeneity and instability were found to dominate the
%systematic error. To reduce the external magnetic fields several types
%of shielding such as active and passive shielding are essential for
%such precision measurements. In addition, internal coils that are
%usually coupled to the passive shields are used to generate the $B_0$
%field. As a result, a detailed understanding of passive and active
%magnetic shielding, magnetic field generation within shielded volumes,
%and precision magnetometry is expected to be crucial to achieve the
%systematic error goals for the next generation of experiments. Much of
%the research and development efforts for these experiments are focused
%on careful design and testing of various magnetic shield geometries
%with precision magnetometers~\cite{altarev2014magnetically,
%  brys2005magnetic,afach2014dynamic,patton2014all}.

%In nEDM experiments, the spin precession frequency $\omega$ of neutrons
%placed in static magnetic $B_0$ and electric $E$ fields is
%measured~(see Chapter~\ref{chap:nedm}). The measured frequencies for
%parallel $\omega_+$ and antiparallel $\omega_-$ relative orientations of the
%fields is sensitive to the neutron electric dipole moment $d_n$
%\begin{equation}
%\hbar\omega_\pm=2\mu_nB_0\pm 2d_nE
%\end{equation}
%where $\mu_n$ is the magnetic moment of the neutron.
%In such experiments, typically $B_0=1~\mu$T is used to provide the
%quantization axis for the UCN. The $B_0$ magnetic
%field generation system typically includes a coil placed within a
%passively magnetically shielded volume.


%Mechanical and temperature changes of the passive magnetic
%shielding~\cite{voigt2013,thiel2007demagnetization}, and the degaussing
%procedure~\cite{thiel2007demagnetization,altarev2015minimizing,sun2016dynamic}
%(also known as demagnetization, equilibration, or idealization),
%affect the stability of the magnetic field within magnetically
%shielded rooms. Active stabilization of the background magnetic field
%surrounding magnetically shielded rooms can also improve the internal
%stability~\cite{voigt2013,afach2014dynamic,Franke_thesis}. The
%current supplied to the $B_0$ coil is generated by an ultra-stable
%current source~\cite{brys2005magnetic}. The coil must also be
%stabilized mechanically relative to the magnetic shielding.

%One additional effect, which is the subject of this chapter, relates to
%the fact that the $B_0$ coil in most nEDM experiments is magnetically
%coupled to the innermost magnetic shield. If the magnetic properties
%of the innermost magnetic shield change as a function of time, it then
%results in a source of instability of $B_0$.  In the present work, we
%estimate this effect and characterize one possible source of
%instability: changes of the magnetic permeability $\mu$ of the
%material with temperature.



The chapter proceeds in the following fashion:
\begin{itemize}
\item The dependence of the internal field on magnetic permeability of
  the innermost shielding layer for a typical nEDM experiment geometry
  is estimated using a combination of analytical and finite element
  analysis techniques.  This sets a scale for the stability
  requirement.
\item New measurements of the temperature dependence of the magnetic
  permeability are presented.  The measurements were done in two ways
  in order to study a variety of systematic effects that were
  encountered.
\item Finally, the results of the calculations and measurements are
  combined to provide a range of temperature sensitivities that takes
  into account sample-to-sample and measurement-to-measurement
  variations.
\end{itemize}

\input{calculations.tex}


\section{Measurements of $\mu(T)$\label{sec:tdep}}

\subsection{Previous Measurements and their Relationship to nEDM
  Experiments\label{sec:previousmeasurement}}

Previous measurements of the temperature dependence of the magnetic
properties of high permeability alloys have been summarized in
Refs.~\cite{couderchon1982some,bozorth1993ferromagnetism,pfeifer1980soft}. These
measurements are normally conducted using a sample of the material to
create a toroidal core, where a thin layer of the material is used in
order to avoid eddy-current and skin-depth
effects~\cite{pfeifer1980soft,kruppvdm}. A value of $\mu$ is
determined by dividing the amplitude of the sensed $B_m$-field by the
amplitude of the driving AC $H_m$-field (similar to the method
described in Section~\ref{sec:transformer}). Normally the frequency
of the $H_m$-field is 50 or 60~Hz.  The value of $\mu$ is then quoted
either at or near its maximum attainable value by adjusting the
amplitude of $H_m$.  Depending on the details of the $B_m-H_m$ curve
for the material in question, this normally means that $\mu$ is quoted
for the amplitude of $H_m$ being at or near the coercivity of the
material~\cite{couderchon1982some,kruppvdm}, resulting in large values
up to $\mu_r=4\times 10^5$.

It is well known that $\mu$ measured in this fashion for toroidal,
thin metal wound cores depends on the annealing process used for the
core.  There is a particularly strong dependence on the take-out or
tempering temperature after the high-temperature portion of the
annealing process has been
completed~\cite{pfeifer1980soft,kruppvdm,couderchon1982some}. Such
studies normally suggest a take-out temperature of 490-500$^\circ$C.
This ensures that the large $\mu_r=4\times 10^{5}$ is furthermore
maximal at room temperature. Slight variations around room
temperature, and assuming the take-out temperature is not controlled
to better than a degree, imply a scale of possible temperature
variation of $\mu$ of approximately
$\left|\frac{1}{\mu}\frac{d\mu}{dT}\right|\simeq 0.3$-1\%/K at room
temperature~\cite{couderchon1982some,kruppvdm}.

A challenge in applying these results to temperature stability of nEDM
experiments is that, when used as DC magnetic shielding, the
high-permeability alloys are usually operated for significantly
different parameters ($B_m$, $H_m$, and frequencies).

For example, when used in a shielding configuration, the effective
permeability is often measured to be typically $\mu_r=20,000$ rather
than $4\times 10^5$. This arises in part because $H_m$ is well below
the DC coercivity.  As noted in Section~\ref{sec:calculation}, a more
appropriate $H_m$ for the innermost magnetic shield of an nEDM
experiment is $<0.007$~A/m, whereas the coercivity is
$H_c=0.4$~A/m~\cite{kruppvdm}.  The frequency dependence of the
measurements could also be an issue.  Typically, nEDM experiments are
concerned with slow drifts at $<0.01$~Hz timescales whereas the
previously reported $\mu(T)$ measurements are performed in an AC mode
at 50-60~Hz. The goal of the experiments discussed in this chapter
was to develop techniques to characterize the material properties of
our own magnetic shields post-annealing, in regimes more relevant to
nEDM experiments.

%-40,000$ for
%3.5~$\mu$T applied external axial and transverse fields at frequencies
%$<0.1$~Hz, where the uncertainty in $\mu_r$ comes a model uncertainty
%resulting from incomplete inclusion of the geometry.



The magnetic shielding factors of each of our prototype four
cylindrical shells, and of various combinations of them, were measured
and found to be consistent with $\mu_r\sim 20,000$.  In our studies of
the material properties of these magnetic shields, two different
approaches to measure $\mu(T)$ were pursued.  Both approaches involved
experiments done using witness cylinders made of the same material and
annealed at the same time as the prototype magnetic shields.  We
therefore expect they have the same magnetic properties as the larger
prototype shields, and they have the advantage of being smaller and
easier to perform measurements with.

The two techniques employed to determine $\mu(T)$ were the following:
\begin{enumerate}
\item measuring the low-frequency AC axial magnetic shielding factor
  of the witness cylinder as a function of temperature, and
\item measuring the temperature dependence of the slope of a minor $B-H$
  loop, using the witness cylinder as a transformer core, similar to
  previous measurements of the temperature dependence of $\mu$, but
  for parameters closer to those encountered in nEDM experiments.
\end{enumerate}
The details and results of each technique are presented below.

% Axial shielding factor measurements Section moved to axial.tex

\input{axial.tex}

% Transformer core measurements Section moved to transformer.tex

\input{transformer.tex}

\section{Relationship to nEDM experiments\label{sec:relationship}}
%\subsection{Summary of $\mu(T)$}

Neutron EDM experiments are typically designed with the DC coil
magnetically coupled to the innermost magnetic shield. As discussed in
Section~\ref{sec:calculation}, if the magnetic permeability of the
shield changes, this results in a change in the field in the
measurement region by an amount
$\frac{\mu}{B_0}\frac{dB_0}{d\mu}=0.01$.

The temperature dependence of $\mu$ has been constrained by two
different techniques using open-ended mu-metal witness cylinders
annealed at the same time as our prototype magnetic shields.  We
summarize the overall result as
0.0\%/K~$<\frac{1}{\mu}\frac{d\mu}{dT}<$~2.7\%/K, where the range is
driven in part by material properties of the different mu-metal
cylinders, and in part by day-to-day fluctuations in the temperature
slopes.



We note the following caveats in relating this measurement to nEDM
experiments:
\begin{itemize}
\item Although the measurement techniques rely on considerably larger
  frequencies and different $H_m$-fields than those relevant to
  typical nEDM experiments, we think it reasonable to assume the
  temperature dependence of the effective permeability should be of
  similar scale.  For frequency, both techniques typically used a 1~Hz
  AC field, whereas for nEDM experiments the field is DC and stable at
  the 0.01~Hz level.  Furthermore, in one measurement technique the
  amplitude of $H_m$ was $\sim 0.004$~A/m and in the other was $\sim
  0.1$~A/m.  For nEDM experiments $H_m<0.007$~A/m and is DC.
\item Both measurement techniques extract an effective $\mu$ that
  describes the slope of minor loops in $B_m-H_m$ space.  A more
  detailed treatment would include a more comprehensive account of
  hysteresis in the material, which is beyond the scope of this work.
\end{itemize}

Our measurement of
0.0\%/K~$<\frac{1}{\mu}\frac{d\mu}{dT}<2.7$\%/K and the generic EDM
experiment sensitivity of $\frac{\mu}{B_0}\frac{dB_0}{d\mu}=0.01$
give a temperature dependence of the magnetic field in a typical
nEDM experiment of $\frac{dB_0}{dT}=0-270$~pT/K.  To achieve a goal of
$\sim 1$~pT stability in the internal field for nEDM experiments, the
temperature of the innermost magnetic shield in the nEDM experiment
should then be controlled to the $<0.004$~K level if the worst-case
dependence is to be taken into account.  This represents a potentially
challenging design constraint for future nEDM experiments.

As noted by others~\cite{khriplovich1997cp}, the use of
self-shielded coils to reduce the coupling of the $B_0$ coil to the
innermost magnetic shield is an attractive option for EDM experiments.
The principle of this technique is to have a second coil structure
between the inner coil and the shield, such that the net magnetic
field generated by the two coils is uniform internally but greatly
reduced externally.  For a perfect self-shielded coil, the field at
the position of the magnetic shield would be zero, resulting in
perfect decoupling, which is to say a reaction factor that is
identically unity.  For ideal geometries, such as spherical
coils~\cite{brown1945fluxball, wheeler1958spherical,purcell1989} or infinitely long
sine-phi coils~\cite{bethBNL,bethUSpatent,bidinosti2005}, the
functional form of the inner and outer current distributions are the
same, albeit with appropriately scaled magnitudes and opposite sign.
More sophisticated analytical and numerical methods have been used
extensively in NMR and MRI to design self-shielded
gradient~\cite{turner1986,hidalgo},
shim~\cite{brideson,forbes}, and transmit
coils~\cite{bidinosti2005,kuzmin}, and should be of value in the
context of nEDM experiments, as well.


\section{Conclusion}

In the axial shielding factor measurement, we found
0.6\%/K~$<\frac{1}{\mu}\frac{d\mu}{dT}<2.7\%$/K, with the measurement
being conducted with a typical $H_m$-amplitude of 0.004~A/m and at a
frequency of 1~Hz.  In the transformer core case, we found
0.0\%/K~$<\frac{1}{\mu}\frac{d\mu}{dT}<2.2\%$/K, with the measurement
being conducted with a typical $H_m$-amplitude of 0.1~A/m and at a
frequency of 1~Hz.

The primary caveat to these measurements is that both measurements
(transformer core and axial shielding factor) do not truly measure
$\mu$.  Rather they measure observables related to the slope of minor
hysteresis loops in $B_m-H_m$ space.  They would be more appropriately
described by a hysteresis model like that of
Jiles~\cite{jiles1994frequency}, but to extract the temperature
dependence of all the parameters of the model is beyond the scope of
this work.  Instead we acknowledge this fact and relate the
temperature dependence of the effective $\mu$ measured by each
experiment.

We think it is interesting and useful information that the two
experiments measure the same scale and sign of the temperature
dependence of their respective effective $\mu$'s.  This is a principal
contribution of this work.

In future work, we plan to measure $B_0(T)$ directly for nEDM-like
geometries using precision atomic magnetometers.  We anticipat based
on the present work that self-shielded coil geometries will achieve
the best time and temperature stability. The R\&D towards
self-shielded coils so far has resulted in three summer student
projects and an honours thesis~\cite{Rosie_thesis}.

%\textbf{add sth about Rosie's stuff here, waiting for Russ's
%  response}.







%\section{Shield Coupled Coils}
%\subsection{Spherical Geometry}
%\subsection{Cylindrical Geometry}


%\section{Method one: Measurement with (How did I used to call it???) }

%\section{Method two: Measurements with a transformer}




%\begin{description}
%\item{This chapter I think should be very similar to the paper. It
%  should be just the extended version of that with more detail (not a
%  lot more though! It means I should not include every single data
%  that I took!)}
%  \item{Motivation: Why are we interested in the temperature
%    dependence of mu. The idea of coupling the field of an internal
%    coil to the shield and that is because the shield act as a return
%    yoke. Then it makes sense to have a stable shield so that the
%    internal field does not change. And then explain that one of the
%    factors are actually the changes in the temperature that affects
%    the magnetic permeability.}
%  \item{The main part of this has to be about the measurements,
%    explain the technique, show several picture of the setup. Look at
%    the long version of the paper on github to remember what we did
%    and why and then include the data that made sense.}
%  \item{Another relevant part is to make the connection between these
%    measurements and the actual changes in the $B$ field and for that
%    I have to include the simulations. An explanation of the OPERA
%    simulations and FEMM. Explain what each one is and what I did.}
%\end{description}


