\chapter{Conclusion\label{chap:overall}}

%Finding a non-zero neutron EDM confirms beyond the standard model
%theories that provide extra sources of CP violation.

The work presented in this thesis is part of the R\&D studies towards
the future nEDM experiment at TRIUMF. The existence of a non-zero nEDM
would indicate a new source of $CP$ violation beyond the standard
model. Theories of new physics beyond the standard model predict such
sources. Based on the Sakharov conditions, new sources of $CP$
violation are essential to create the observed baryon asymmetry in the
universe.

The focus of the research in this thesis was on the two aspects of the
nEDM measurement:
\begin{itemize}
\item Magnetic field stability~(focused on temperature related
  effects)
\item UCN production and storage
\end{itemize}
and I first summarize these results.
\section{ Magnetic Field Requirements and Stability}
To measure the nEDM, an ensemble of polarized UCN are placed in the
presence of aligned electric and magnetic fields. The Larmor
precession frequency of the UCN is measured once when the electric and
magnetic fields are parallel, and once when they are
anti-parallel. The frequency shift between these two geometries is
proportional to the nEDM. In this process, the existence of a very
stable and homogeneous magnetic environment is essential. The applied
DC magnetic field should be held constant. To achieve the magnetic
requirements several layers of magnetic shielding are employed
including active and passive shielding. Interal coils are placed
inside the passive shielding to create the DC magnetic field.  In the
past, experiments used coils where the return yoke corresponded to the
innermost magnetic shielding layer. These are referred to as
``shield-coupled'' coils. In this case, a change in the properties of
the innermost passive shield, such as its magnetic permeability $\mu$,
would affect the magnetic field measured internally. One of the
factors that can cause such changes in $\mu$ is the changes in the
environmental temperature.

%\section{Overview of Temperature Dependence of Mangetic Permeability}

In Chapter~\ref{chap:muofT} the result of the studies of the changes
of $\mu$ with temperature were presented and discussed. Two methods
were pursued to study the correlation of the changes in the measured
internal magnetic field with respect to the changes in temperature.
For those measurements, we used small witness cylinders~(15.2~cm in
length and 5.2~cm in diameter) made of the same material and annealed
at the same time as the prototype passive shields. As a result, we
expected them to have the same magnetic properties of the larger
passive shields while having the advantage of being easier to perform
experiments with.


In method one, a witness cylinder was put inside a coil system that
produced a low-frequency magnetic field. The axial shielding factor as
a function of temperature was then measured. These measurements were
repeated with two different coils to study the systematic effects. In
the second technique, which is more common, the witness cylinder was
used as a core of a transformer. A primary and a secondary coil were
wound on the witness cylinder. Here the slopes of the minor $B-H$
loops as a function of temperature were measured.

The measurements were conducted in AC fields with frequencies around
1~Hz as opposed to the DC fields in the actual nEDM experiments. To
relat the shielding factor measurements to $\mu(T)$, finite element
simulations were performed to find the shielding factor of the witness
cylinders as a function of $\mu$. Combining the measurements and the
simulations, it was found that
0.6\%/K~$<\frac{1}{\mu}\frac{d\mu}{dT}<2.7\%$/K with an
$H_m$-amplitude of 0.004~A/m at 1~Hz. In the second method, it was
found that 0.0\%/K~$<\frac{1}{\mu}\frac{d\mu}{dT}<2.2\%$/K with a
typical $H_m$-amplitude of 0.1~A/m at 1~Hz. For a typical nEDM
experiment, $H_m < 0.007$~A/m.

Considering the overall range implied by the measurements of
0.0\%/K~$<\frac{1}{\mu}\frac{d\mu}{dT}<2.7$\%/K and the generic EDM
experiment sensitivity of $\frac{\mu}{B_0}\frac{dB_0}{d\mu}=0.01$, the
temperature dependence of the magnetic field in a typical nEDM
experiment would be $\frac{dB_0}{dT}=0-270$~pT/K. This means, to
safely achieve the magnetic stability goal of 1~pT in the interal
field~(for the worst case temperature slope), the temperature of the
innermost magnetic shield in the nEDM experiment should be controlled
to $<0.004$~K level which puts a challenging constraint on the future
nEDM experiment design. Achieving this level of temperature control is
very difficult. As a result, TUCAN collaboration is now considering
the self-shielded coil designs as the baseline for the future nEDM
experiment.

\section{Overview of UCN Studies at TRIUMF}

The second half of this thesis was focused on the new UCN facility at
TRIUMF. In 2016, the prototype vertical UCN source, previously built
and tested in Japan, was shipped to TRIUMF. The unique feature of this
facility is the production UCN by combining spallation neutrons with a
superfluid helium converter.

In November 2017 the first UCN were produced with the prototype
vertical source at TRIUMF. The experiments that were conducted include
measurements of the UCN yield, UCN storage lifetime and UCN rate in
the steady-state mode of operation.  Those experiments and their
simulations were presented in Chapter~\ref{chap:UCNresult}. Such
experiments are essential for a better understanding of the UCN
cryostat and for the design of the next generation UCN source.

Around 40,000 UCN were detected for the standard measurement of
1~$\mu$A proton beam current while irradiating the target for
60~s. The maximum number of UCN produced was 325,000 at 10~$\mu$A
proton beam current. Over the three weeks of the experimental run
period, the measured storage lifetime of UCN dropped from 37~s to
27~s. This is likely due to contamination of the source when the UCN
valve was opened. The UCN yield for the standard measurements at
1~$\mu$A proton beam current and 60~s irradiation time also dropped
correspondingly by about 40\%. Other than the source contamination,
different experimental configuration in the second half of the
experimental period caused this drop.

We could successfully produce three times more UCN than ever
before. However, our UCN storage lifetime was only half of the
previously best storage lifetime. This is most likely due to a
combination of the contamination in the source, the existence of the
burst disk, and the non-optimized UCN valve for the UCN storage. The
result of the simulations proved that the upscattering rate in the
superfluid helium follows $\tau^{-1}_{\mathrm He} = B T^7$. In
addition, the UCN yield and the UCN storage lifetime simulations
matched the experimental values best when the $B$ values were between
0.008/s and 0.016/s. At higher temperatures above 1~K, the
upscattering in the helium vapour plays a significant role.

%Some measurements with the heater tapes were also performed to find
%the amount of heat load on the cryostat when applying the proton
%beam. Even though the result of those experiments was not conclusive,
%it gave valuable insight for the future experimental plan.


\section{Overall Conclusion and Future Work}
The research presented in this thesis is part of the TUCAN's ongoing
R\&D towards the future nEDM measurement at TRIUMF. The experimental
goal is to measure the nEDM with the sensitivity level of
$10^{-27}$~e$\cdot$cm.

The best previous measurement of the nEDM showed that the dominant
systematic uncertainty is due to the magnetic field instability. As a
result, one source of systematic uncertainty, the temperature
dependence of magnetic permeability $\mu$, was studied. The result of
those measurements as well as the relevant simulations showed that, to
fulfill our 1~pT internal magnetic field stability requirement, the
temperature of the innermost magnetic shield for the nEDM experiment
should be controlled to better than 4~mK level while employing the
shield coupled coils to generate the DC field.

The current nEDM experiments suffer from low UCN statistics. The UCN
experiments with our prototype UCN source provide the prerequisites
for building our next-generation UCN source with cooling power and UCN
flux increased by two orders of magnitude~(see
Section~\ref{sec:triumfnedm}). The excellent match of the simulations
and experiments presented in Chapter~\ref{chap:UCNresult} makes us
confident that we can predict the performance of this future source
and experiment very well.




%The TUCAN team is focused on the R\&D for the next generation UCN
%source.







